Species persistence under climate and fishing

### 2014-01-29 Writing/MS.tex
cleaned up conflicts

When the climate changes, the area in which organisms can survive and reproduce moves through space. 
This change does not occur in isolation but rather appears on a background of other disturbances. We used an 
integrodifference model explicitly accounting for dispersal, reproduction, a shifting environment, and 
harvesting to examine how two disturbances, range shift and harvesting, interact and govern population 
persistence. We found threshold rates of harvesting and of environmental shift that are needed to allow the 
population to persist, studied how these critical parameters depend on the growth rate and dispersal behavior 
of the population, and measured the interactions between the stressors. In particular, we found low but positive 
synergy between the two stressors: harvesting aggravates the population's sensitivity to a shifting range. 
Finally, we introduced two conservation techniques into simulations of the population model -- threshold 
harvest rules and marine protected areas (MPAs) -- and found that 
these approaches could mitigate the negative interaction of the two stressors.  
\hspace{10cm}

@@

\noindent Second, we examined harvesting rules more complex than harvesting a constant proportion of the population. Whereas population persistence in the analytical model does not depend on the functional form of recruitment $f$, we must specify a recruitment function to perform simulations.  We chose to use a Beverton-Holt function,

@@

The equilibrium biomass of the population decreases as either the harvesting pressure 
increases or the environmental shifts more quickly (Figure \ref{baseline}). If the equilibrium biomass is $0$, this 
indicates that the harvesting pressure has exceeded the critical harvesting rate $h^*$ and the environmental 
is shifting more quickly than the critical rate of environmental shift $c^*$. As the harvesting rate $h$ increases, 
the critical rate of environmental shift $c^*$ decreases: the environment must move more slowly to 
accommodate the population growing more slowly (Figure \ref{baseline}). Conversely, as the rate of 
environmental shift $c$ increases, the critical harvesting rate $h^*$ decreases (Figure  \ref{baseline}). This 
means that a harvesting rate that is sustainable in the absence of environmental shift may no longer be sustainable if 
the environment starts shifting. The simulations replicate the analytical results with the critical speed $c^*$ 
declining as the critical harvest rate $h^*$ increases and vice versa.

@@

\begin{figure}[htbp]
\begin{subfigure}{3in}
\subcaption{\label{biomass}}
\includegraphics[width=\textwidth]{plots/eqbiomass.pdf}
\end{subfigure}
\begin{subfigure}{3in}
\subcaption{\label{rates}}
\includegraphics[width=\textwidth]{plots/critical_rates.pdf}
\end{subfigure}
\caption{
%(a) The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis. These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.  (b) The critical harvesting rate on the y-axis as a function of the rate of environmental shift on the x-axis.  Black lines correspond to a growth rate of $R_0=3$, red to $R_0=7$, and blue to $R_0=10$.  Solid lines correspond to an average dispersal distance $\langle d \rangle =0.1$ and dashed lines correspond to an average dispersal distance $\langle d \rangle =0.25$.  These results are from an approximated Gaussian dispersal kernel with $L=1$.
}
\label{baseline}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=3in]{plots/synergy.pdf}
\caption{
%Positive synergy between the two stressors.  The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.  These results are from an approximated Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
}
\label{Synergy}
\end{center}
\end{figure}

@@

\begin{subfigure}{.33\textwidth}
\subcaption{}
\includegraphics[width=\textwidth]{plots/eqbiomass_sim.pdf}
\end{subfigure}
\begin{subfigure}{.33\textwidth}
\subcaption{}
\includegraphics[width=\textwidth]{plots/eqbiomass_thresh.pdf}
\end{subfigure}
\begin{subfigure}{.33\textwidth}
\subcaption{}
\includegraphics[width=\textwidth]{plots/eqbiomass_mpa.pdf}
\end{subfigure}
\caption{
%The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.  (a) No management.  (b) Threshold harvesting levels.  (c) MPAs.  These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$.
}
\label{management}
\end{figure}

@@

Figure \ref{management}: The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.  (a) No management.  (b) Threshold harvesting levels.  (c) MPAs.  These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$.

>>>>

When the climate changes, the area in which organisms can survive and reproduce moves through space. 
This change does not occur in isolation but rather appears on a background of other disturbances. We used an 
integrodifference model explicitly accounting for dispersal, reproduction, a shifting environment, and 
harvesting to examine how two disturbances, range shift and harvesting, interact and govern population 
persistence. We found threshold rates of harvesting and of environmental shift such that more extreme harvesting or shift will drive the population extinct and studied how these critical parameters depend on the growth rate and dispersal behavior 
of the population.  We then measured the interaction between the stressors and we found low but positive 
synergy between the two stressors: harvesting aggravates the population's sensitivity to a shifting range. 
Finally, we introduced two conservation techniques into simulations of the population model -- threshold 
harvest rules and marine protected areas (MPAs) -- and found that 
these approaches could mitigate the negative interaction of the two stressors.  
\hspace{10cm}

@@

\noindent Second, we examined harvesting rules more complex than harvesting a constant proportion of the population. Whereas population persistence in the analytical model does not depend on the functional form of recruitment $f$, to perform simulations we must 
specify a recruitment function.  Again, we chose to use a Beverton-Holt function.

@@

As the harvesting rate $h$ increases, 
the critical rate of environmental shift $c^*$ decreases: the environment must move more slowly to 
accommodate the population growing more slowly (Figure \ref{baseline}). Conversely, as the rate of 
environmental shift $c$ increases, the critical harvesting rate $h^*$ decreases (Figure  \ref{baseline}). This 
means that a harvesting rate that is sustainable in the absence of environmental shift may no longer be sustainable if 
the environment starts shifting.    When the harvesting pressure has exceeded the critical harvesting rate $h^*$ or the environmental 
is shifting more quickly than the critical rate of environmental shift $c^*$, the biomass of the population at equilibrium will be equal to $0$.  Before those thresholds are reached, the equilibrium biomass of the population decreases as either the harvesting pressure 
increases or the environmental shifts more quickly (Figure \ref{baseline}). The simulations replicate the analytical results with the critical speed $c^*$ 
declining as the critical harvest rate $h^*$ increases and vice versa.

@@

\begin{figure}[htbp]
\begin{subfigure}{3in}
\subcaption{\label{rates}}
\includegraphics[width=\textwidth]{plots/critical_rates.pdf}
\end{subfigure}
\begin{subfigure}{3in}
\subcaption{\label{biomass}}
\includegraphics[width=\textwidth]{plots/eqbiomass.pdf}
\end{subfigure}

\caption{(a) The critical harvesting rate on the y-axis as a function of the rate of environmental shift on the x-axis.  Black lines correspond to a growth rate of $R_0=3$, red to $R_0=7$, and blue to $R_0=10$.  Solid lines correspond to an average dispersal distance $\langle d \rangle =0.1$ and dashed lines correspond to an average dispersal distance $\langle d \rangle =0.25$.  (b) The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis. These results are from a Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.    These results are from an approximated Gaussian dispersal kernel with $L=1$.}

\label{baseline}
\end{figure}

@@

\begin{figure}[htbp]
\begin{center}
\includegraphics[width=3in]{plots/synergy.pdf}
\caption{
Positive synergy between the two stressors.  The x-axis shows the rate of environmental shift, the y-axis shows the harvesting rate, and the color indicates the loss in biomass in the doubly stressed population in excess of the sum of the losses caused by each stressor individually, $E_\text{hc}-E_\text{h}-E_\text{c}$.  These results are from an approximated Gaussian dispersal kernel with parameters $L=1$, $R_0=5$, $\langle d \rangle = 0.399$.
}
\label{Synergy}
\end{center}
\end{figure}

@@

\begin{subfigure}{.33\textwidth}
\subcaption{}
\includegraphics[width=\textwidth]{plots/eqbiomass_sim.pdf}
\end{subfigure}
\begin{subfigure}{.33\textwidth}
\subcaption{}
\includegraphics[width=\textwidth]{plots/eqbiomass_thresh.pdf}
\end{subfigure}
\begin{subfigure}{.33\textwidth}
\subcaption{}
\includegraphics[width=\textwidth]{plots/eqbiomass_mpa.pdf}
\end{subfigure}
\caption{
The equilibrium biomass of the population as a function of the rate of environmental shift on the x-axis and the harvesting rate on the y-axis with and without management strategies.  (a) No management.  (b) Threshold harvesting levels.  (c) MPAs.  These results are from a simulation with a Laplacian dispersal kernel with parameters $L=1$, $R_0=5$, $K=100$, and $\langle d \rangle =2$.
}
\label{management}
\end{figure}

@@

\section{Discussion}
We extended a spatial model to study a fish population subject to both the climate-driven range shifts and harvesting. Previous work has highlighted the importance of reproduction and dispersal to the vulnerability of fish populations to climate change \citep[\citet{Fordhametal2013}]{Hastingsetal2005} and our model explicitly includes these processes.  In particular, we derived expressions for critical harvesting rates and critical rates of environmental shift for a separable Gaussian kernel and found these critical rates using numerical simulations of a Laplace kernel.  For each kernel we studied, we found that the higher the growth rate and the better the mean dispersal distance matches the rate of environmental shift, the better a population can adjust to harvest and climate change.  More interestingly, we found a negative relationship between the critical harvesting rate and the rate of environmental shift.  That is, the more quickly the environment shifts the less harvesting it takes to drive the population extinct.  The curved line separating parameters that will allow the population to persist from those that won't is on indication of an interaction between the stressors.

The two management strategies we modeled -- harvesting thresholds and MPAs -- both increased the population's biomass at equilibrium and improved its ability to persist.  Protected areas have been advanced as a way to help organisms keep pace with range shifts, as well as to ameliorate anthropogenic disturbances like harvesting and habitat fragmentation\citep{Lawleretal2010, Hannahetal2007,Botsfordetal2001, Gaylordetal2005, HastingsBotsford2003,Thomasetal2012}.  Our results, that spatial management increased the maximum harvesting rate and maximum climate velocity at which the population could survive, corroborate these previous findings.  The management strategies also affect how the two stressors interact with each other.  Specifically, harvesting thresholds can alleviate interactions between the two stressors.   While the management strategies only change harvesting practices and do not directly address affect climate change, understanding how they ameliorate synergistic affects between harvesting and range shifts will help to better implement and place protected areas.
